---
title: "Data Collection and Data Wrangling"
author: "Ghulam Nabi"
date: "April 18, 2021"
output: html_document
---

##### NOTE: 
1. Change author name and date to your exercise submission date in above section
2. Your code MUST execute without errors. 
3. You can add more lines in your code as required.

## Section 1: Manipulating Data

### Question 1 
**The dataset is loaded for you. Perform the following tasks:**  
1. use the USPersonalExpenditure dataset and store in a variable.  
2. Compute summary statistics: mean, standard deviation (sd) across time for each of the indicators.  
3. Create a new column that contains average values for each of the indicators.  


```{r }
library(help='datasets')
```


```{r }
USPersonalExpenditure
```

```{r }
us_personal <- USPersonalExpenditure
us_personal <- as.data.frame(us_personal)
us_personal
```


```{r }
library(dplyr)
sum_arise <- summarise_at(us_personal,vars("1940","1945","1950","1955","1960"),funs(mean,sd),na.rm=T)
sum_arise
mutation <- mutate(us_personal,average_value = summarise_at(us_personal,vars("1940","1945","1950","1955","1960"),funs(mean),na.rm=T))
mutation


```


```{r }

```


```{r }

```


```{r }

```

### Question 2 
**download the data from the available URL:**  
1. Create a new column containing the average bodywt (body weight) of each genus and of each order.  
2. Create a dataframe containing the average REM sleep time for each order.  
3. How many missing values are there (in total) and per column?  
4. How would you like to impute missing values? Write Justification.  <i> Hint: Overall Mean/media/mode vs. Groupby Mean/media/mode?


```{r }
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/msleep_ggplot2.csv"
filename <- "msleep_ggplot2.csv"
```


```{r }
download.file(url,destfile = "msleep_ggplot2.csv")
```


```{r }
msleep_ggplot2 <- read.csv("msleep_ggplot2.csv")
msleep_ggplot2
```


```{r }
mutate(msleep_ggplot2,average_bodywt = summarise(msleep_ggplot2,mean=mean(bodywt),na.rm=T)) %>%
  select(genus,order,average_bodywt)


```


```{r }
mutate(msleep_ggplot2,average_rem=summarise(msleep_ggplot2,mean=mean(sleep_rem),na.rm=T)) %>% select(average_rem,order)
```


```{r }
sum(is.na(msleep_ggplot2))
# there are total 136 missing values 
colSums(is.na(msleep_ggplot2)) # per column

# Justification of imputing missing values 
# I would completely take mean/mode/median of all values as find it more easy and all values will be fit in with easy way
```

### Good job! You've completed this section!

## Section 2: Tidyverse

### Question 1
**Use the above dataset and perform the following tasks using any library from tidyverse:**  
1. Filter results to print average REM sleep and average total sleep  for those animals who are carnivores and then for those who are primates.  
2. Use the order column and "spread" it across the rest of the observations.  


```{r }
filter(msleep_ggplot2,order==c("carnivores","primates")) %>% select(order,sleep_rem,sleep_total)
```


```{r }
library(tidyr)
spread(msleep_ggplot2,order,name)
spread(msleep_ggplot2,order,genus)
spread(msleep_ggplot2,order,vore)
spread(msleep_ggplot2,order,conservation)
spread(msleep_ggplot2,order,sleep_total)
spread(msleep_ggplot2,order,sleep_rem)
spread(msleep_ggplot2,order,sleep_cycle)
spread(msleep_ggplot2,order,awake)
spread(msleep_ggplot2,order,brainwt)
spread(msleep_ggplot2,order,bodywt)
```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```

### Good job! You've completed this entire exercise!
